---
title: "p8105_hw6_xl3371"
author: "Xuer Liu"
date: "2023-12-02"
output: github_document
---

```{r setup, include = FALSE}
library(tidyverse)
library(purrr)
library(broom)
library(ggplot2)
library(modelr)

knitr::opts_chunk$set(
  echo = TRUE, 
  warning = FALSE,
  fig.width = 6, 
  fig.asp = .6, 
  out.width = "90%"
  )

theme_set(theme_minimal() + theme(legend.position = "bottom"))
```

## Problem 2
```{r, message=FALSE}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2022-01-01",
    date_max = "2022-12-31") |>
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) |>
  select(name, id, everything())
```

```{r}
set.seed(5)

# Function to generate a bootstrap sample
boot_sample <- function(df) {
  sample_frac(df, replace = TRUE)
}

# Create 5000 bootstrap samples
boot_straps <- tibble(strap_number = 1:5000) %>%
  mutate(strap_sample = map(strap_number, \(i) boot_sample(weather_df)))

fit_model <- function(df) {
  model <- lm(tmax ~ tmin + prcp, data = df)
  r_squared <- broom::glance(model)$r.squared
  beta_product <- broom::tidy(model)$estimate[2] * broom::tidy(model)$estimate[3]
  return(list(r_squared = r_squared, beta_product = beta_product))
}

```


```{r}
# Apply function to each sample and gather results
boot_results <- boot_straps %>%
  mutate(metrics = map(strap_sample, fit_model)) %>%
  unnest_wider(metrics)

```

```{r}
# Function to plot density
plot_density <- function(data, column, title) {
  ggplot(data, aes(x = .data[[column]])) +
    geom_density() +
    labs(title = title)
}

# Plot distributions
plot_density(boot_results, "r_squared", "Distribution of R-squared Estimates")
plot_density(boot_results, "beta_product", "Distribution of log(beta1 * beta2) Estimates")

```

The density curve for $\hat r^2$ is unimodal and slightly left skewed distribution with a peak around 0.90. The density curve for $\log(\hat\beta_1\times \hat\beta_2)$ is also unimodal but has a more complex shape with what appears to be two peaks, one near -0.01 and another small peak near 0.00. 

```{r}
# calculate and print CIs
calculate_and_print_ci <- function(data, column) {
  ci <- data %>%
    filter(!is.nan(!!sym(column))) %>%
    summarize(
      ci_lower = quantile(!!sym(column), 0.025),
      ci_upper = quantile(!!sym(column), 0.975)
    ) %>%
    knitr::kable(digits = 3)
  print(ci)
}

calculate_and_print_ci(boot_results, "r_squared")
calculate_and_print_ci(boot_results, "beta_product")

```

## Problem 3
```{r, message=FALSE}
birthweight_df <- 
  read_csv("./data/birthweight.csv", na = c("", "NA", "Unknown")) %>% 
  janitor::clean_names() %>%
  mutate(
    babysex = as.factor(babysex),
    frace = as.factor(frace),
    malform = as.factor(malform),
    mrace = as.factor(mrace)
  ) 

# check for missing data
sum(is.na(birthweight_df))
```


In my model, I will include `gaweeks`, `blength`, `bhead`, `smoken`, and `ppbmi` with the reasons based on a hypothesized structure for the factors that underly birthweight:

* Gestational age in weeks (`gaweeks`): the number of weeks they spend developing before birth is closely linked to their weight at birth.
* Baby’s length at birth (`blength`): longer babies have had more time to grow and tend to weigh more.
* Baby’s head circumference at birth (`bhead`): a larger head circumference can correlate with a higher birth weight, as it indicates good brain and skull development, which contributes significantly to total birth weight.
* Average number of cigarettes smoked per day during pregnancy (`smoken`): smoking during pregnancy can lead to lower birth weight babies.
* Mother’s pre-pregnancy BMI (`ppbmi`): A mother's pre-pregnancy body mass index is indicative of her nutritional and health status before pregnancy, which can impact fetal growth.

```{r}
my_model <- lm(bwt ~ gaweeks + blength + bhead + smoken + ppbmi, data=birthweight_df)

summary(my_model)
```

Based on summary, all estimates I chose are significant.


```{r, message=FALSE}
birthweight_df %>% 
  add_predictions(my_model) %>% 
  add_residuals(my_model) %>% 
  ggplot(aes( x = pred, y = resid)) + 
  geom_point(alpha = 0.5) +
  geom_smooth() +
  labs(
    title = "Model residuals against fitted values",
    x = "Fitted Values",
    y = "Residuals"
      )
```

```{r}
library(caret)

# Model 1: main effect
model1 <- lm(bwt ~ blength + gaweeks, data=birthweight_df)

# Model 2: three-way interaction
model2 <- lm(bwt ~ bhead * blength * babysex, data=birthweight_df)

# Function to calculate cross-validated prediction error
cv_error <- function(my_model, data) {
  cv <- trainControl(method = "cv", number = 10)
  train(bwt ~ ., data = data, method = "lm", trControl = cv)$results$RMSE
}

# Cross-validated prediction errors
error_original <- cv_error(my_model, birthweight_df)
error_model1 <- cv_error(model1, birthweight_df)
error_model2 <- cv_error(model2, birthweight_df)

# Comparison
error_comparison <- data.frame(Original = error_original, Model1 = error_model1, Model2 = error_model2)

error_comparison

```

Model 1, which is the one using length at birth and gestational age as predictors (main effects only), has the lowest RMSE, which suggests that it is the best model for predicting birthweight among the three models, based on the cross-validated prediction error.
